Supplementary Material for the Paper ``Enriching Object Detection with\\2D-3D Registration and Continuous Viewpoint Estimation''

### 2014-11-18 approach.tex
typo and capion4
minor change

Whitening the non-zero
feature vector $x'$ means a solving the system of linear equations, $\Sigma' w' = (x' -
\mu')$. In
\cite{Hariharan12}, the authors make use of the fact that covariance matrices
are symmetric and positive semidefinite to solve the system via the Cholesky
decomposition with Gaussian Elemination, which requires $O(n^3)$ time.

@@

\begin{figure}[t]
  \begin{center}
  \begin{tabular}{cc}
     \includegraphics[width=0.5\linewidth]{whotime} & 
     \includegraphics[width=0.5\linewidth]{speedup}\\
     (a) & (b) \\
 \end{tabular}
  \end{center}
  \caption{(a) Runtime analysis of whitening. HOG means feature
    extraction time, WHO-Chol uses ~\cite{Hariharan12} and
    WHO-CG uses iterative Conjugate Gradient. NZ-WHO-CG uses only
    non-zero cells and Conjugate Gradient. (b) final speedup
    of NZ-WHO vs. WHO.}
  \label{fig:whotime}
\end{figure}
%
\begin{figure}[t]
  \centering
  \includegraphics[width=0.5\linewidth]{residual}
  \caption{Residuals from different method}
  \label{fig:whoresidual}
\end{figure}

>>>>

Whitening the non-zero
feature vector $x'$ means a solving the system of linear equations, $\Sigma' w' = (x' -
\mu')$. In
\cite{Hariharan12}, the authors make use of the fact that covariance matrices
are symmetric and positive semidefinite to solve the system via the Cholesky
decomposition with Gaussian Elimination, which requires $O(n^3)$ time.

@@

\begin{figure}[t]
  \begin{center}
  \begin{tabular}{cc}
     \includegraphics[width=0.5\linewidth]{whotime} & 
     \includegraphics[width=0.5\linewidth]{speedup}\\
     (a) & (b) \\
 \end{tabular}
  \end{center}
  \caption{(a) Runtime analysis of whitening. HOG means feature
    extraction time, WHO-Chol uses ~\cite{Hariharan12} and
    WHO-CG uses iterative Conjugate Gradient. NZ-WHO-CG (Ours) uses only
    non-zero cells and Conjugate Gradient. (b) final speedup
    of NZ-WHO vs. WHO.}
  \label{fig:whotime}
\end{figure}
%
\begin{figure}[t]
  \centering
  \includegraphics[width=0.5\linewidth]{residual}
  \caption{Residuals from different method}
  \label{fig:whoresidual}
\end{figure}

### 2014-11-18 egpaper_final.tex
minor change

We propose an efficient method for synthesizing templates from 3D models that
runs on the fly -- that is, it quickly produces detectors for an arbitrary
viewpoint of a 3D model without expensive dataset-dependent training or template
storage. Given a 3D model and an arbitrary continuous detection viewpoint, our
method synthesizes a discriminative template by extracting features from a
rendered view of the object and decorrelating spatial dependences among the
features. Our decorrelation procedure relies on a gradient-based algorithm that
is more numerically stable than standard decomposition-based procedures, and we
efficiently search for candidate detections by computing FFT-based template
convolutions. Due to the speed of our template synthesis procedure, we are able
to perform joint optimization on continuous scale, translation, rotation, and
focal length. We provide an efficient GPU implementation of our algorithm, and
we validate its performance on 3D Object Classes and PASCAL3D+ datasets.

>>>>

We propose an efficient method for synthesizing templates from 3D models that
runs on the fly -- that is, it quickly produces detectors for an arbitrary
viewpoint of a 3D model without expensive dataset-dependent training or template
storage. Given a 3D model and an arbitrary continuous detection viewpoint, our
method synthesizes a discriminative template by extracting features from a
rendered view of the object and decorrelating spatial dependences among the
features. Our decorrelation procedure relies on a gradient-based algorithm that
is more numerically stable than standard decomposition-based procedures, and we
efficiently search for candidate detections by computing FFT-based template
convolutions. Due to the speed of our template synthesis procedure, we are able
to perform joint optimization on scale, translation, continuous rotation, and
focal length. We provide an efficient GPU implementation of our algorithm, and
we validate its performance on 3D Object Classes and PASCAL3D+ datasets.

### 2014-11-18 egpaper_supp.tex
minor change

More qualitative results are on the Fig.~\ref{fig:3dobject_fig}. The each of the template in the ensemble of NZ-WHO templates has a corresponding rendering and a CAD model. These can be transferred as a metadata and overlay the rendering on top of the detection.

\begin{figure*}[h]
\setlength\tabcolsep{1pt}
\centering
\begin{tabular}{|c|c|}
  \hline
  \includegraphics[width=0.40\linewidth]{supp/car1.png} &
  \includegraphics[width=0.40\linewidth]{supp/bicycle1.png} \\
  \includegraphics[width=0.40\linewidth]{supp/car23.png} &
  \includegraphics[width=0.40\linewidth]{supp/bicycle3.png} \\
  \includegraphics[width=0.40\linewidth]{supp/car26.png} &
  \includegraphics[width=0.40\linewidth]{supp/bicycle4.png} \\
  \includegraphics[width=0.40\linewidth]{supp/car29.png} & 
  \includegraphics[width=0.40\linewidth]{supp/bicycle5.png} \\
  \includegraphics[width=0.40\linewidth]{supp/car10.png} &
  \includegraphics[width=0.40\linewidth]{supp/bicycle9.png} \\
  \includegraphics[width=0.40\linewidth]{supp/car13.png} &
  \includegraphics[width=0.40\linewidth]{supp/bicycle13.png} \\
  \includegraphics[width=0.40\linewidth]{supp/car15.png} &
  \includegraphics[width=0.40\linewidth]{supp/bicycle11.png} \\
  \includegraphics[width=0.40\linewidth]{supp/car8.png} &
  \includegraphics[width=0.40\linewidth]{supp/bicycle12.png} \\
  \includegraphics[width=0.40\linewidth]{supp/car20.png} & 
  \includegraphics[width=0.40\linewidth]{supp/bicycle15.png} \\
  \hline
  Car & Bicycle \\
  \hline
  \end{tabular}
\caption{Detection results on 3D Object dataset car and bike classes. In each column, original image (left) and detection result overlaid on top (right).}% with a bounding box and corresponding confidence score (right).}
  \label{fig:3dobject_fig}
\end{figure*}

>>>>

More qualitative results are on the Fig.~\ref{fig:3dobject_good} and Fig.~\ref{fig:3dobject_bad}. The each of the template in the ensemble of NZ-WHO templates has a corresponding rendering and a CAD model. These can be transferred as a metadata and overlay the rendering on top of the detection.

\begin{figure*}[h]
\setlength\tabcolsep{1pt}
\centering
\begin{tabular}{|c|c|}
  \hline
  \includegraphics[width=0.40\linewidth]{supp/car32.png} &
  \includegraphics[width=0.40\linewidth]{supp/bicycle17.png} \\
  \includegraphics[width=0.40\linewidth]{supp/car26.png} &
  \includegraphics[width=0.40\linewidth]{supp/bicycle18.png} \\
  \includegraphics[width=0.40\linewidth]{supp/car29.png} & 
  \includegraphics[width=0.40\linewidth]{supp/bicycle13.png} \\
  \includegraphics[width=0.40\linewidth]{supp/car10.png} &
  \includegraphics[width=0.40\linewidth]{supp/bicycle9.png} \\
  \includegraphics[width=0.40\linewidth]{supp/car1.png} &
  \includegraphics[width=0.40\linewidth]{supp/bicycle11.png} \\
  \includegraphics[width=0.40\linewidth]{supp/car8.png} &
  \includegraphics[width=0.40\linewidth]{supp/bicycle12.png} \\
  \includegraphics[width=0.40\linewidth]{supp/car20.png} & 
  \includegraphics[width=0.40\linewidth]{supp/bicycle15.png} \\
  \includegraphics[width=0.40\linewidth]{supp/car31.png} & 
  \includegraphics[width=0.40\linewidth]{supp/bicycle10.png} \\
  \includegraphics[width=0.40\linewidth]{supp/car33.png} & 
  \includegraphics[width=0.40\linewidth]{supp/bicycle16.png} \\
  \hline
  Car & Bicycle \\
  \hline
  \end{tabular}
\caption{Successful detection results on 3D Object dataset car and bike classes. In each column, original image (left) and detection result overlaid on top (right).}% with a bounding box and corresponding confidence score (right).}
  \label{fig:3dobject_good}
\end{figure*}

\begin{figure*}[h]
\setlength\tabcolsep{1pt}
\centering
\begin{tabular}{|c|c|}
  \hline
  \includegraphics[width=0.40\linewidth]{supp/car13.png} &
  \includegraphics[width=0.40\linewidth]{supp/bicycle1.png} \\
  \includegraphics[width=0.40\linewidth]{supp/car30.png} &
  \includegraphics[width=0.40\linewidth]{supp/bicycle5.png} \\
  \includegraphics[width=0.40\linewidth]{supp/car18.png} &
  \includegraphics[width=0.40\linewidth]{supp/bicycle19.png} \\
  \hline
  Car & Bicycle \\
  \hline
  \end{tabular}
\caption{Failed pose estimation results on 3D Object dataset car and bike classes. In each column, original image (left) and detection result overlaid on top (right).}% with a bounding box and corresponding confidence score (right).}
  \label{fig:3dobject_bad}
\end{figure*}

### 2014-11-18 experiments.tex
minor change

To that end, we create an ensemble of NZ-WHO templates (Sect.~\ref{sec:approach})
using $9$ different CAD models and a total of 192 different
viewpoints, using 4 different evelation angles, 24 azimuth angles,
and 2 focal lengths. We run the entire ensemble exhaustively over
each test image, in a sliding window fashion.

@@

This result is encouraging, since our approach reaches a level of
performance that is on par to current state-of-the-art while at the
same time being much faster to train. It takes merely a few minutes to
train, while both ALM~\cite{Xiang12} and DPM-VOC+VP~\cite{Pepik12} are
complex models that optimize non-convex objective functions during
training, which is only made tractable by resorting to delayed
constraint generation in the form of hard negative mining, and can
easily take a day on a single machine. In addition, our approach uses
only rendered images, avoiding the need for real-world training data
with costly bounding box and viewpoint annotations.
%
\begin{table}[!htbp]
    \footnotesize
  \begin{center}
    \begin{tabular}{|c|c|c|c|}
    \hline
     AP/MPPE& Ours & ALM\cite{Xiang12} & VOC-DPM\cite{Pepik12} \\
    \hline\hline
    car & \textbf{99.8}/91.7 &  98.4/93.4 & \textbf{99.8}/\textbf{97.5} \\ 
    bicycle & 93.0/87.4 & 93.0/91.4 & \textbf{98.8}/\textbf{97.5} \\
    \hline
    \end{tabular}
  \end{center}
  \caption{Average Precision (AP) and Mean Precision in Pose
    Estimation (MPPE) on 3D Object Classes~\cite{Savarese07} cars.}% Our method produces high quality 2D-3D matching yet performs on par with state-of-the art detectors.}
  \label{tab:3dobject}
\end{table}

@@

For both cases, we compare the performance of a discrete incarnation
of our method (Sect.~\ref{sec:approach}) and our full model, including the
fine-tuning based on MCMC~\cite{Sect.~\ref{sec:tuning}}

Quantitative results are given in Tab.~\ref{tab:pascal12} in terms of
AP and AVP and in Fig.~\ref{fig:car_cnn_ap} in terms of
precision-recall plots and viewpoint confusion matrices. Example
outputs of our method applied to candidate object detections are given
in Fig.~\ref{fig:pascal12cnn}, and the effect of fine-tuning is
visualized in Fig.~\ref{fig:tuning}.

 More qualitative results are part of
supplemental material

\paragraph{Results.}
%
In Tab.~\ref{tab:pascal12}, we make mainly two observation. First, we
see that adding our method to DPM-VOC+VP consistently improves
performance in terms of AVP, for both car and bicycle, and across all
viewpoint bins. This is already the case for our discrete method: the
improvement ranges from $0.3$ for bicycle-8v to $5.8$ for car-4v. 
%
Using the R-CNN as the base detector increases AVP even more, in
particular for the bicycle class: for bicycle-4v, our discrete method
improves over the corresponding DPM-VOC+VP result by $12.6$.

>>>>

To that end, we create an ensemble of NZ-WHO templates (Sect.~\ref{sec:approach})
using $9$ different CAD models and a total of 192 different
viewpoints, using 4 different elevation angles, 24 azimuth angles,
and 2 focal lengths. We run the entire ensemble exhaustively over
each test image, in a sliding window fashion.

@@

This result is encouraging, since our approach reaches a level of
performance that is on par to current state-of-the-art while at the
same time being much faster to train. It takes merely a few minutes to
train, while both ALM~\cite{Xiang12} and DPM-VOC+VP~\cite{Pepik12} are
complex models that optimize non-convex objective functions during
training, which is only made tractable by resorting to delayed
constraint generation in the form of hard negative mining, and can
easily take a day on a single machine. In addition, our approach uses
only rendered images, avoiding the need for real-world training data
with costly bounding box and viewpoint annotations.
%
\begin{table}[!htbp]
    \footnotesize
  \begin{center}
    \begin{tabular}{|c|c|c|c|}
    \hline
     AP/MPPE& Ours & ALM\cite{Xiang12} & VOC-DPM\cite{Pepik12} \\
    \hline\hline
    car     & \textbf{99.8} / 91.7 & 98.4 / 93.4 & \textbf{99.8} / \textbf{97.5} \\ 
    bicycle & 93.0 / 90.9          & 93.0 / 91.4 & \textbf{98.8} / \textbf{97.5} \\
    \hline
    \end{tabular}
  \end{center}
  \caption{Average Precision (AP) and Mean Precision in Pose
    Estimation (MPPE) on 3D Object Classes~\cite{Savarese07} cars.}% Our method produces high quality 2D-3D matching yet performs on par with state-of-the art detectors.}
  \label{tab:3dobject}
\end{table}

@@

For both cases, we compare the performance of a discrete incarnation
of our method (Sect.~\ref{sec:approach}) and our full model, including the
fine-tuning based on MCMC (Sect.~\ref{sec:fine}).

Quantitative results are given in Tab.~\ref{tab:pascal12} in terms of
AP and AVP and in Fig.~\ref{fig:car_cnn_ap} in terms of
precision-recall plots and viewpoint confusion matrices. Example
outputs of our method applied to candidate object detections are given
in Fig.~\ref{fig:pascal12cnn}, and the effect of fine-tuning is
visualized in Fig.~\ref{fig:tuning}. More qualitative results are part of
supplemental material
\paragraph{Results.}
%
In Tab.~\ref{tab:pascal12}, we make mainly two observation. First, we
see that adding our method to DPM-VOC+VP consistently improves
performance in terms of AVP, for both car and bicycle, and across all
viewpoint bins. This is already the case for our discrete method: the
improvement ranges from $0.3$ for bicycle-8v to $5.8$ for car-4v. 
%
Using the R-CNN as the base detector increases AVP even more, in
particular for the bicycle class: for bicycle-4v, our discrete method
improves over the corresponding DPM-VOC+VP result by $12.6$.

### 2014-11-18 mcmc.tex
added back
minor change

We model the probability that the object in the test image
$\mathcal{I}$ is generated by the parameters $p$ as a distribution in
the exponential family, and let
\begin{align}
    P(p) & \sim e^{ \max_{s} w(p) \ast \mathcal{T}_s(\mathcal{I})},
\end{align}
where $\max_{s} w(p) \ast T_s(\mathcal{I})$ is the maximum convolution score of
NZ-WHO template $w(p)$ with image feautures $T_s(\mathcal{I})$ for all scale $s$, as defined
in Sect.~\ref{sec:nzwho}.

@@

This algorithm changes the current state $p$ to a new state $q$ based
on the acceptance probability
\begin{align}
    A(p \rightarrow q) & =  min\left( 1,  \frac{P(q) g(q \rightarrow p)}{P(p) g(p \rightarrow q)}\right).
\end{align}
We define $3$ different types of moves that can alter the state
(Fig.~\ref{fig:moves}), {\em (i)} changing the focal length $f$, {\em
(ii)} one of the rotational pose parameters $v_i$, and {\em (iii)} CAD model
index $m$. For (i) and (ii), we use Gaussian proposal distributions, and a uniform
distribution over model changes for (iii). For scale and translation, we
implicitly compute all possible translation and scale by convolving $w(q)$ to
HOG pyramid.
\begin{align}
    g(p \rightarrow p + v_i') & \sim \mathcal{N}(\mu = 0,\sigma = \sigma_v) \quad i \in \{1,2,3\}\\
    g(p \rightarrow p + f') & \sim \mathcal{N}(\mu = 0,\sigma = \sigma_f)\\
    g(p \rightarrow p + m') & \sim c \delta(m) + (1-c) Unif(1,M),
\end{align}
where $g$ is the proposal distribution. % and $P$ is the distribution of which we
%defined using Gaussian distribution around previous state. $A$ is the
%acceptance probability where we accept new proposal with $A(x \rightarrow x')$.
%
% One can also think of the image as an image cropped by detection
% bounding box. We synthesize covariance matrix and create NZ-WHO
% template $w(x)$ for particular viewpoint, focal length and CAD model.
% In sum, for new proposal $x'$, we make a NZ-WHO template on the fly 
% We found out that it requires good initialization to converge to
% local minima.

@@

\paragraph{Initialization.}
Since the Metropolis-Hastings algorithm is susceptible to getting
stuck in local optima, we need a good initialiation in order to
increase the probability to find a solution that is close to the
global optima. We achieve this by first running a discrete,
pre-trained version of our algorithm (i.e., an ensemble of NZ-WHO
templates) in order to get promising candidate 2D bounding boxes and
poses to start from. We then initialize $x_0$ for each of these candidates.

>>>>

We model the probability that the object in the test image
$\mathcal{I}$ is generated by the parameters $p$ as a distribution in
the exponential family, and let
\begin{align}
    P(p) & \sim e^{ \max_{s} w(p) \ast \mathcal{T}_s(\mathcal{I})},
\end{align}
where $\max_{s} w(p) \ast T_s(\mathcal{I})$ is the maximum convolution score of
NZ-WHO template $w(p)$ with image features $T_s(\mathcal{I})$ for all scale $s$, as defined
in Sect.~\ref{sec:nzwho}.

@@

This algorithm changes the current state $p$ to a new state $q$ based
on the acceptance probability
\begin{align}
    A(p \rightarrow q) & =  min\left( 1,  \frac{P(q) g(q \rightarrow p)}{P(p) g(p \rightarrow q)}\right).
\end{align}
We define $3$ different types of moves that can alter the state
(Fig.~\ref{fig:moves}), {\em (i)} changing the focal length $f$, {\em
(ii)} one of the rotational pose parameters $v_i$, and {\em (iii)} CAD model
index $m$. For (i) and (ii), we use Gaussian proposal distributions, and a uniform
distribution over model changes for (iii). For scale and translation, we
implicitly compute all possible translation and scale by convolving $w(q)$ to
HOG pyramid.
\begin{align}
    g(p \rightarrow p + v_i') & \sim \mathcal{N}(\mu = p_{v_i},\sigma = \sigma_v) \quad i \in \{1,2,3\}\\
    g(p \rightarrow p + f') & \sim \mathcal{N}(\mu = p_{f},\sigma = \sigma_f)\\
    g(p \rightarrow p + m') & \sim c \delta(p_m) + (1-c) Unif(1,M),
\end{align}
where $g$ is the proposal distribution. % and $P$ is the distribution of which we
%defined using Gaussian distribution around previous state. $A$ is the
%acceptance probability where we accept new proposal with $A(x \rightarrow x')$.
%
% One can also think of the image as an image cropped by detection
% bounding box. We synthesize covariance matrix and create NZ-WHO
% template $w(x)$ for particular viewpoint, focal length and CAD model.
% In sum, for new proposal $x'$, we make a NZ-WHO template on the fly 
% We found out that it requires good initialization to converge to
% local minima.

@@

\paragraph{Initialization.}
Since the Metropolis-Hastings algorithm is susceptible to getting
stuck in local optima, we need a good initialization in order to
increase the probability to find a solution that is close to the
global optima. We achieve this by first running a discrete,
pre-trained version of our algorithm (i.e., an ensemble of NZ-WHO
templates) in order to get promising candidate 2D bounding boxes and
poses to start from. We then initialize $x_0$ for each of these candidates.
